Natural Selection Across Three Generations of Americans

### 2024-02-16 response-to-reviewers-revision-1.qmd

*Finally, the statement ’in the UK Biobank sample, correlations of PGS coefficients between younger and older parents are negative (-0.319), and for 42 out of 58 PGS, coefficients have opposite signs. However, in bootstraps, these statistics are not significantly different from 0 and 29 respectively.’ was not entirely clear to me. What does it mean that 42 PGS had opposite sign effects? Does that mean for old versus younger parents’ coefficients were very different, even in the opposite direction? How does this relate to the first part of the sentence that coefficients were negative. Also, why has this been tested against 0 and 29?*

The earlier claim is not "coefficients are negative overall" but 
"coefficients among younger parents are negatively correlated with coefficients
among older parents". We are dealing with correlations of correlations: this
can get confusing.

Negatively correlated, and more strongly, opposite-signed, effects between
young and old parents support the theory that income effects dominate for
older parents while substitution effects dominate for younger parents. To
be a bit more explicit: if a PGS simply doesn't correlate with human capital,
we expect it to not be selected on among either group of parents. If a PGS
is correlated with human capital, that will lead to income and substitution effects
for those with more human capital. If the substitution effect dominates for
younger parents then the PGS will have a negative correlation with fertility;
if the income effect dominates for older parents, then the PGS will have a positive
correlation with fertility.

The test against 29 had the following logic: if all effects were really zero,
and estimates were pure noise with mean 0, then we'd expect opposite-signed and same-signed "effects" to be equally likely; so, 29 is half of the 58 scores.

However, thinking about this issue a bit more made us change our tests more
generally. The original situation was that we tested most pairs of coefficients
across sample subgroups by comparing their absolute size. Coefficients for
older and younger parents had similar absolute sizes, but we noticed they were
opposite-signed and hence ran the tests above. This is not very satisfactory: it's 
too like special pleading, and it also suggests the original tests weren't adequate.

We now improve our tests in table 1 as follows. We take the null hypothesis
to be "selection coefficients are the same in the different subgroups, for all
PGIs". Under the null, the true difference between selection coefficients is zero; the 
difference in our sample (and in bootstrapped subsamples) should then not
be significantly different from zero. Our alternative hypothesis is that
selection coefficients are smaller among the richer/more educated/older/married
group. That is, if the coefficient among the poorer group is $x_p$, then

$x_r \times sign(x_p) < x_p \times sign(x_p)$

Note that, unlike tests of absolute values, the above always holds if 
$x_r$ and $x_p$ are opposite signed.

Under the null, we'd expect the above equation to hold for just 50%
of the scores. If it holds for significantly more than 50% of the scores
(using bootstrapped C.I.s) then we can reject the null in favour of the 
alternative.

We also try a simpler alternative approach. Using the few PGIs with the
largest and most precisely estimated selection effects in the whole sample,
we estimate selection effects in different subgroups and test for a significant
difference between them, via a simple interaction term. We report these results
too.

TE TODO: the above^^^

@@

> The results here support prediction 1 but are more ambiguous for prediction 2. PGS coefficients on RLRS also appear larger for low-income groups and unmarried respondents. But there is little evidence for larger coefficients among people with lower education, or younger parents. This may be due to the low sample size. But in the UK, the between-group differences were large; differences that big would surely have been visible here. The theory can accommodate this non-result, if preferences for income are less curved in the US for whatever reason. But note that *any* theory with a negative relationship between education and fertility will give prediction 1. In this sense, results here are less supportive of the economic theory specifically.

We don't think these results give a simple thumbs up or down to the theory,
but we're open to suggestions that would help clarify what we say.

*The division into a "black" and a "white" is necessary due to technical reasons, but further complicates the results/discussion. Presumably the results for the larger sample (I think in the case of the HRS this are the  whites) should be presented in the main text, and the results for the smaller population moved into a supplement?*

We're OK with this, since the black sample is really too small to say much.

@@

We now discuss some health-related traits in the conclusion:

> Whilst past research has focused on [cognitive] traits, we find selection 
differentials are often of similar magnitudes for health related traits. The 
selection differential for self-rated health is greater than those of education 
and cognitive performance. The most significant, positively selected trait was 
ADHD, which was also found in the UK [@hugh2022human]. Future research should 
study the health and medical implications of natural selection, in addition to 
their social implications.

*What is the data base for the mean number of children of people born in the same year - is pooling appropriate, as the means may vary considerably according to ethnicity?*

We thought about this, and agree that it is better to use mean number of 
children of whites and of blacks separately for each ethnic group.

Neither option is perfect. If we calculate RLRS separately for whites and
blacks, we are effectively treating them as separate breeding populations.
Black and white people had low rates of intermarriage in the 20th century but
that is a fact of law and culture, not of biology. Conversely, if we calculate
RLRS by taking the mean number of children across the whole population, we
end up effectively making claims about natural selection in the US population
as a whole, by taking the white population as representative - which it very
likely isn't.

Overall we prefer the former option, so we have now done so and added a brief
discussion. Effectively, the white US population is a "single data point" for
theory testing. (Even if we could pool the whole US population, that is still
not the whole human population; this point is not much discussed in the literature.)
